---
title: "R Assignment #3"
author: "Put Your Name Here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the packages we'll need
library(tidyverse)
library(lubridate)
library(readxl)
library(writexl)
library(janitor)

```

## House Democrats - First Trump Impeachment

### Load the data

Run the code below to load the data.

It will create a dataframe called `impeach`, which contains a row for every House Democrat and whether or not the member publicly called for impeachment in the case of the *first* Trump impeachment.

Additionally, a series of election results and demographics are included related to each member's district.

```{r}

impeach <- readRDS("impeach.rds")


```

### Questions

Write your code using grouping in the chunks below to help answer the following questions.

1)  How many members in the data set favor impeachment vs. not (using the `for_impeachment` column)?
Here, I grouped by for impeach as instructed and then summarized count contributions to ge the totals for those who voted yes vs. no. I found that 209 members favor impeachment while 26 members do not. 

```{r}

impeach %>%
  group_by(for_impeachment) %>% 
  summarise(count_contribs = n())




```

2)  Similar to #1 above, using grouping to now break down the Democratic House members by TWO measures: those who are for or against impeachment (as you did above)....and then districts above/below national GDP.  
  
You'll want to ensure you do the grouping on both columns together, e.g. `group_by(column1, column2)`

Here, I filtered democrats since I'm only looking at democratic house members, I then group by for impeachmen and gdp above national to get that 7 who voted no are from districts above national GDP, 19 are below, and 126 who voted yes are above with 83 below.

```{r}

impeach %>%
  group_by(for_impeachment, gdp_above_national) %>%
summarise(count_contribs = n())

  








```

3)  Now do the same as #2, but this time instead of GDP, group by whether the district is above or below the national average for the percentage of college graduates. The column that captures this information is `pct_bachelors_compared_to_national`.

Here, I pasted the code from question 2 but replaced gdp_above_national with pct_bachelors_compared_to_national.

```{r}


impeach %>%
  filter(party == "D") %>%
  group_by(for_impeachment, pct_bachelors_compared_to_national) %>%
summarise(count_contribs = n())







```

4)  Let's look at the college graduation comparison in a slightly different way.  
  
Instead of counting how many districts are above/below the national average, this time summarize by the MEAN percentage of college grads (located in the column `pct_bachelors`) for districts that are Yes for impeachment vs. No.  
  
In other words, you want to end up with the calculated mean for what that percentage is for the Yes districts and the No districts.

Here, I grouped by for impeachment and summarized the mean of pct bachelors to find the average of college graduates, and got approximately 27% for no and 33% for yes.

```{r}

impeach %>%
  group_by(for_impeachment) %>%
  summarise(mean(pct_bachelors))






```

5)  Do the same as #4, but this time show the MEAN *percentage of the vote that Donald Trump received* for districts that are Yes for impeachment vs. No.  
The relevant column for that is `trump_percent`.

Here, I grouped by for impeachment and summarized the mean and got approximately 43% for no and 32% for yes.

```{r}


impeach %>%
  group_by(for_impeachment) %>%
  summarise(mean(trump_percent))







```

6)  Filter out only the members who are a yes for impeachment. Then of those "Yes" member, how many won their 2018 election by less than 5 percentage points (`margin_flag_2018`) vs. more?
I filtered the impeachment votes to only show me members that voted yes and then grouped the margin flag and summarized to get that 17 won by 5 points or less and 192 won with more than 5 points.

```{r}

  
impeach %>%
  filter(for_impeachment == "YES") %>%
  group_by(margin_flag_2018) %>%
  summarise(count = n())





```

7)  Come up with another breakdown of your choosing for how you'd like to examine this dataset. Say what you'll look at, and then put the code below to find the answer.
Here, I wanted to determine which state had the most people vote yes for impeachment and found that CA had the most.

```{r}

  
impeach %>%
  filter(for_impeachment == "YES") %>%
  group_by(state) %>%
  summarise(count = n()) %>%
arrange(desc(count))





```

7)  You'll see a column in the table called date_announced. For members who came out as a Yes for impeachment, this is the date they announced their support for it.  
  
Use the mutate() function to create two new columns: one that extracts the year from date_announced, and a second that extracts the month.

Here I used mutate to make a year announced and month announced column by specifying that each column consists of either the month or year from the date announced data.

```{r}


  impeach %>%
  mutate (year_announced = year(date_announced), 
month_announced = month(date_announced))
  







```

8) Using the new columns you created in #7, use grouping to count up how many House Democrats during each month announced their support of impeachment.
Here I pasted the same mutation code from above and then grouped by month and summarized to get January-3, April-7, May-39 etc. 

```{r}

impeach %>%
  mutate (year_announced = year(date_announced), 
month_announced = month(date_announced)) %>%
  group_by(month_announced) %>%
  summarise(count = n())







```



